---
title: "Example_with_EUCAST_Data"
author: "Angelika Moder, Michaela Raab, Nicole Halmdienst, Sebastian Schrey, Wolfgang Rannetbauer"

output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_with_EUCAST_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This documents is an introduction to the package `emGauss`. The aim of this package is to find the cutoff point (ECOFF) between resistant and non-resistant antibiotics.

# Setup

Therefore, first load the package `emGauss`.
```{r setup}
library(emGauss)
```

# Data
This package also includes real data (Zonedata) from the EUCAST website. This will be loaded in the next lines of code

```{r}
data("ZD", package = "EUCASTData")
```

# Example Dataset
In order to get an example dataset, we have to extract the nessercary columns from the ZD dataset. The plot shows that the data may have two components.

```{r}
# 65, 855, 1091 example datasets 

zd <- as.numeric(ZD[1091, 4:48])
zd.data <- data.frame(bin = 6:50, nrObs = zd)
barplot(zd.data$nrObs, names.arg = c(zd.data$bin))

```


# Find optimum number of components
In this section we use the function `em.gauss.opti.groups` to find the optimum number of components based on the BIC. 
```{r}
k <- 5
y <- as.numeric(zd)

result <-em.gauss.opti.groups(y = y,
                     k = k,
                     alpha = 3,
                     beta= 1,
                     method = "quantile",
                     epsilon=0.0001)
```
As a result we get a list with all parameters of the fitted distribution and the AIC and BIC values.
The next line of code shows the best number of components based on the BIC. As we can see is the optimum number of components 2.
```{r}
plot(result$BIC, ylab = 'BIC',
     xlab = 'Number of components')
abline(v=which.min(result$BIC), lty= 2, col=2)

```


# Find start values

In order to find start values for the mean and the variance we use the function `createCluster`.


```{r}
k <- 2
start.musigma2 <- createCluster(as.matrix(zd.data), k, 
                                method = 'quantile')
start.musigma2

```


# Fitting distribution and find ECOFF
After finding the start values we can use the function `em.gauss` to fit the mixing distribution of normals to the data. As a result we get the mean, variance, mixing proportions and loglikelihood of the fitted distribution. Furthermore the function provides an estimate of the ECOFF value.
```{r}
y <- as.numeric(zd)

em.result <- em.gauss(y = y,
                      mu = start.musigma2$mu,
                      sigma2 = start.musigma2$sigma2,
                      pi = rep(1/k, k),
                      alpha = 3,
                      beta = 1,
                      epsilon = 0.0001,
                      max.iter = 650)
```

# Plot result
Now we use the function plot.fct to plot the fitted distribution and the ECOFF.

```{r}
plot_fct(y, 
          mu = em.result$mu, 
          sigma2 = em.result$sigma2, 
          pi = em.result$pi, 
          ecoff = em.result$ecoff)
```


